Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach

### 2018-12-07 Manuscript.Rmd
Change to exploratory ECDF of absolute FV magnitudes so that all plots are plotted along the same column for comparison

library(igraph)
library(pander)
library(ggnetwork)
library(ggplot2)
library(gridExtra)
library(data.table)
library(plyr)
library(mcclust)
library(lme4)
library(parallel)
library(corrplot)
library(DescTools)
library(knitr)

@@

# Permutation analysis
# Note: this doesn't care whether a region was uni or bilaterally activated. Maybe do that.
# Load the coordinates as studies in a list, because for the contrast the studies are permuted
# Use table(unlist(list)) to get the counts in that case
setwd('./DEC')
templist <- dir()
studyList_sv <- lapply(templist, readr::read_csv, col_names = F)

setwd('../DMN/')
templist <- dir()
studyList_dmn <- lapply(templist, readr::read_csv, col_names = F)

# For each permutation, reassign the study labels, maintaining the original number of studies per condition 
allStudies <- c(studyList_sv, studyList_dmn)
nPerms <- 5000
nullDist <- numeric()

for (i in seq(nPerms)) {

  if(i %in% c(nPerms*0.25, nPerms*0.5, nPerms*0.75, nPerms*0.95)) {print(paste("Completed", i/nPerms, "percent of the permutation..."))}

  # Sample all the studies and divide into 2 groups
  randPerm <- sample(allStudies)
  tmpSV <- randPerm[seq(length(studyList_sv))]
  tmpDMN <- randPerm[-seq(length(studyList_sv))]

  # Count the number of times each region was reported in the literature
  # Bilateral activations still count as 1, so we might want to reconsider that
  tmpDF <- data.frame(Parcel = gsub("_","", Yeo_labels),
                      SV = sapply(Yeo_labels, function(roi) {length(grep(roi, tmpSV))}),
                      DMN = sapply(Yeo_labels, function(roi) {length(grep(roi, tmpDMN))}))

  # Perform the prop test for each pair and get the max chi-squared stat
  nullDist[i] <- max(
                  apply(tmpDF[,2:3], 1, function(data) {
                    prop.test(data, c(length(tmpSV), length(tmpDMN)))$statistic
                    }
                  ), na.rm = T)

# Now do the same for the actual data, to see how it compares to the null dist
tmpDF <- data.frame(Parcel = gsub("_","", Yeo_labels),
                    SV = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_sv))}),
                    DMN = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_dmn))}),
                    propSV = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_sv))}) / length(studyList_sv),
                    propDMN = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_dmn))}) / length(studyList_dmn))

# Perform the prop test for each pair and get the max chi-squared stat + pvalue based on the computed null distribution
tmpDF$Chi_squared <- apply(tmpDF[,2:3], 1, function(data) {
  round(prop.test(data, c(length(studyList_sv), length(studyList_dmn)))$statistic, digits = 2)
  })
tmpDF$Pval <- 1 - ecdf(nullDist)(tmpDF$Chi_squared) # probability of finding a chi-squared stat at least as large as the observed under a permuted null distribution

# ROIs that survived the threshold based on Chi-squared value (>95th quantile from null distribution -> 8.43)
significantROIs <- tmpDF %>% filter(Chi_squared >= quantile(nullDist,0.95))

@@

``` {r ROI computation, echo=FALSE, message = FALSE}
#TO DO: CHANGE THIS SO IT'S BASED ON STUDYLIST_SV/DMN, COUNTING JUST LIKE IN THE PERMUTATION ABOVE
## ROI definition
# First, load the raw count csv files
valCounts <- read.csv2('decCounts.csv', sep = ",")
dmnCounts <- read.csv2('dmnCounts.csv', sep = ",")

# Then, get the means of the counts per domain, both within/outside the search space
# This is to get a broad estimate of the average report amounts in the literature
tempBoot <- bootstrap(valCounts$Counts)
BV <- round(mean(tempBoot))
BVquant <- round(quantile(tempBoot, c(0.025, 0.975)), digits = 2)
tempBoot <- bootstrap(dmnCounts$Counts)
BD <- round(mean(tempBoot))
BDquant <- round(quantile(tempBoot, c(0.025, 0.975)), digits = 2)
rm(tempBoot)

# Now select only the ROIs from Yeo et al (TEMP)
valCounts <- getCoords(Labels = Yeo_labels, Coords = valCounts)$Coords
dmnCounts <- getCoords(Labels = Yeo_labels, Coords = dmnCounts)$Coords

# Now, for every ROI, do a prop test to check if the proportion of reports is significantly greater
# than the mean report over the number of studies (i.e. sample-based null probability, which could be changed to perm)
for (i in seq(valCounts$Counts)) {
  indx <- valCounts$Counts[i]
  tempTest <- prop.test(indx, 27, alternative = "greater", p = BV/27)
  valCounts$Chi[i] <- tempTest$statistic
  valCounts$pVal[i] <- tempTest$p.value
}

for (i in seq(dmnCounts$Counts)) {
  indx <- dmnCounts$Counts[i]
  tempTest <- prop.test(indx, 77, alternative = "greater", p = BD/77)
  dmnCounts$Chi[i] <- tempTest$statistic
  dmnCounts$pVal[i] <- tempTest$p.value
}

# Retain only the significant parcels
t1 <- valCounts$Label[valCounts$pVal < 0.05]
t2 <- dmnCounts$Label[dmnCounts$pVal < 0.05]

# Remove hemispheric references
t1 <- unique(substring(t1,2))
t2 <- unique(substring(t2,2))

# Labels to put in text
dmnLbls <- gsub("ROI","", t2) 
dmnLbls <- gsub("_","", dmnLbls)
valLbls <- gsub("ROI","", t1) 
valLbls <- gsub("_","", valLbls)

# final overlap
bothLbls <- dmnLbls[dmnLbls %in% valLbls]

@@

``` {r Load Summary Data}
# Load data
setwd('./Summary/')
temp <- list.files()
subjList <- sapply(temp, substring, first=1, last=6)
Summaries <- lapply(temp, read.csv)
```

@@

### Testing grounds
# this part compares the distribution of Fiedler values for DMN and nonDMN
# note: remove filter(y > 0) for 'full brain'
# First plot compares the median FV between communities across subjects
test <- do.call(rbind, Summaries)
test %>%
  mutate(SubjID = rep(subjList, each = 5081),
         FV_abs = abs(FiedlerVec),
         Community = ifelse(FiedlerBinary == 0, "nonDMN", "DMN")) %>%
  filter(y > 0) %>%
  group_by(SubjID, Community) %>%
  summarize(FV_mean = median(FV_abs)) %>%
  ggplot(aes(Community, FV_mean)) +
  geom_boxplot() + 
  theme_classic()
# Second one plots the ecdf
# Line indicates the 75th quantile of the combined Fiedler vectors across participants (i.e. potential threshold)
test %>%
  mutate(SubjID = rep(subjList, each = nrow(Summaries[[1]])),
         FV_abs = abs(FiedlerVec),
         Community = ifelse(FiedlerBinary == 0, "nonDMN", "DMN")) %>%
  filter(y > 0) %>%
  ggplot(aes(FV_abs, group = interaction(SubjID, Community), color = Community)) +
  stat_ecdf() +
  geom_vline(xintercept = 0.009, lty = 2) +
  theme_classic()
```

>>>>

# Libraries used for the whole process
library(tidyverse)
library(igraph)
library(pander)
library(ggnetwork)
library(ggplot2)
library(gridExtra)
library(data.table)
library(plyr)
library(mcclust)
library(lme4)
library(parallel)
library(corrplot)
library(DescTools)
library(knitr)

@@

# # Permutation analysis
# # Note: this doesn't care whether a region was uni or bilaterally activated. Maybe do that.
# # Load the coordinates as studies in a list, because for the contrast the studies are permuted
# # Use table(unlist(list)) to get the counts in that case
# setwd('./DEC')
# templist <- dir()
# studyList_sv <- lapply(templist, readr::read_csv, col_names = F)
# 
# setwd('../DMN/')
# templist <- dir()
# studyList_dmn <- lapply(templist, readr::read_csv, col_names = F)
# 
# # For each permutation, reassign the study labels, maintaining the original number of studies per condition 
# allStudies <- c(studyList_sv, studyList_dmn)
# nPerms <- 5000
# nullDist <- numeric()
# 
# for (i in seq(nPerms)) {
#   
#   if(i %in% c(nPerms*0.25, nPerms*0.5, nPerms*0.75, nPerms*0.95)) {print(paste("Completed", i/nPerms, "percent of the permutation..."))}
#     
#   # Sample all the studies and divide into 2 groups
#   randPerm <- sample(allStudies)
#   tmpSV <- randPerm[seq(length(studyList_sv))]
#   tmpDMN <- randPerm[-seq(length(studyList_sv))]
#   
#   # Count the number of times each region was reported in the literature
#   # Bilateral activations still count as 1, so we might want to reconsider that
#   tmpDF <- data.frame(Parcel = gsub("_","", Yeo_labels),
#                       SV = sapply(Yeo_labels, function(roi) {length(grep(roi, tmpSV))}),
#                       DMN = sapply(Yeo_labels, function(roi) {length(grep(roi, tmpDMN))}))
#   
#   # Perform the prop test for each pair and get the max chi-squared stat
#   nullDist[i] <- max(
#                   apply(tmpDF[,2:3], 1, function(data) {
#                     prop.test(data, c(length(tmpSV), length(tmpDMN)))$statistic
#                     }
#                   ), na.rm = T)
#   
# }
# 
# # Now do the same for the actual data, to see how it compares to the null dist
# tmpDF <- data.frame(Parcel = gsub("_","", Yeo_labels),
#                     SV = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_sv))}),
#                     DMN = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_dmn))}),
#                     propSV = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_sv))}) / length(studyList_sv),
#                     propDMN = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_dmn))}) / length(studyList_dmn))
# 
# # Perform the prop test for each pair and get the max chi-squared stat + pvalue based on the computed null distribution
# tmpDF$Chi_squared <- apply(tmpDF[,2:3], 1, function(data) {
#   round(prop.test(data, c(length(studyList_sv), length(studyList_dmn)))$statistic, digits = 2)
#   })
# tmpDF$Pval <- 1 - ecdf(nullDist)(tmpDF$Chi_squared) # probability of finding a chi-squared stat at least as large as the observed under a permuted null distribution
# 
# # ROIs that survived the threshold based on Chi-squared value (>95th quantile from null distribution -> 8.43)
# significantROIs <- tmpDF %>% filter(Chi_squared >= quantile(nullDist,0.95))

@@

``` {r ROI computation, echo=FALSE, message = FALSE}
# #TO DO: CHANGE THIS SO IT'S BASED ON STUDYLIST_SV/DMN, COUNTING JUST LIKE IN THE PERMUTATION ABOVE
# ## ROI definition
# # First, load the raw count csv files
# valCounts <- read.csv2('decCounts.csv', sep = ",")
# dmnCounts <- read.csv2('dmnCounts.csv', sep = ",")
# 
# # Then, get the means of the counts per domain, both within/outside the search space
# # This is to get a broad estimate of the average report amounts in the literature
# tempBoot <- bootstrap(valCounts$Counts)
# BV <- round(mean(tempBoot))
# BVquant <- round(quantile(tempBoot, c(0.025, 0.975)), digits = 2)
# tempBoot <- bootstrap(dmnCounts$Counts)
# BD <- round(mean(tempBoot))
# BDquant <- round(quantile(tempBoot, c(0.025, 0.975)), digits = 2)
# rm(tempBoot)
# 
# # Now select only the ROIs from Yeo et al (TEMP)
# valCounts <- getCoords(Labels = Yeo_labels, Coords = valCounts)$Coords
# dmnCounts <- getCoords(Labels = Yeo_labels, Coords = dmnCounts)$Coords
# 
# # Now, for every ROI, do a prop test to check if the proportion of reports is significantly greater
# # than the mean report over the number of studies (i.e. sample-based null probability, which could be changed to perm)
# for (i in seq(valCounts$Counts)) {
#   indx <- valCounts$Counts[i]
#   tempTest <- prop.test(indx, 27, alternative = "greater", p = BV/27)
#   valCounts$Chi[i] <- tempTest$statistic
#   valCounts$pVal[i] <- tempTest$p.value
# }
# 
# for (i in seq(dmnCounts$Counts)) {
#   indx <- dmnCounts$Counts[i]
#   tempTest <- prop.test(indx, 77, alternative = "greater", p = BD/77)
#   dmnCounts$Chi[i] <- tempTest$statistic
#   dmnCounts$pVal[i] <- tempTest$p.value
# }
# 
# # Retain only the significant parcels
# t1 <- valCounts$Label[valCounts$pVal < 0.05]
# t2 <- dmnCounts$Label[dmnCounts$pVal < 0.05]
# 
# # Remove hemispheric references
# t1 <- unique(substring(t1,2))
# t2 <- unique(substring(t2,2))
# 
# # Labels to put in text
# dmnLbls <- gsub("ROI","", t2) 
# dmnLbls <- gsub("_","", dmnLbls)
# valLbls <- gsub("ROI","", t1) 
# valLbls <- gsub("_","", valLbls)
# 
# # final overlap
# bothLbls <- dmnLbls[dmnLbls %in% valLbls]

@@

``` {r Load Summary Data}
# Load data
setwd('./Summary/')
temp <- list.files()
subjList <- sapply(temp, substring, first=1, last=6)
Summaries <- lapply(temp, read.csv)
nVertices <- nrow(Summaries[[1]]) # in case I need the number of vertices included per brain
```

@@

### Testing grounds
# this part compares the distribution of Fiedler values for DMN and nonDMN
# note: remove filter(y > 0) for 'full brain'
# First plot compares the median FV between communities across subjects
test <- do.call(rbind, Summaries)
ecdfPCC <- test %>%
          mutate(SubjID = rep(subjList, each = nVertices),
                 FV_abs = abs(FiedlerVec),
                 Community = ifelse(FiedlerBinary == 0, "nonDMN", "DMN")) %>%
          filter(y < 0) %>%
          ggplot(aes(FV_abs, group = interaction(SubjID, Community), color = Community)) +
          stat_ecdf(alpha = 0.3) +
          #scale_color_manual(values = c("Yellow", "Red")) +
          xlim(0,0.05) +
          geom_vline(xintercept = 0.009, lty = 2) +
          labs(title = "Per-subject ECDF of Fiedler values: PCC") +
          theme_classic()

ecdfPFC <- test %>%
          mutate(SubjID = rep(subjList, each = nVertices),
                 FV_abs = abs(FiedlerVec),
                 Community = ifelse(FiedlerBinary == 0, "nonDMN", "DMN")) %>%
          filter(y > 0) %>%
          ggplot(aes(FV_abs, group = interaction(SubjID, Community), color = Community)) +
          #scale_color_manual(values = c("Yellow", "Red")) +
          stat_ecdf(alpha = 0.3) +
          xlim(0,0.05) +
          labs(title = "Per-subject ECDF of Fiedler values: PFC") +
          geom_vline(xintercept = 0.009, lty = 2) +
          theme_classic()

ecdfALL <- test %>%
  mutate(SubjID = rep(subjList, each = nVertices),
         FV_abs = abs(FiedlerVec),
         Community = ifelse(FiedlerBinary == 0, "nonDMN", "DMN")) %>%
  ggplot(aes(FV_abs, group = interaction(SubjID, Community), color = Community)) +
  #scale_color_manual(values = c("Yellow", "Red")) +
  stat_ecdf(alpha = 0.5) +
  xlim(0,0.05) +
  labs(title = "Per-subject ECDF of Fiedler values") +
  geom_vline(xintercept = 0.009, lty = 2) +
  theme_classic()

#grid.arrange(pcc1,pfc1, ncol = 2)
#grid.arrange(All,pcc2,pfc2, ncol = 1)
